---
title: "quarto_demo"
format: html
editor: visual
---

## Deseq2[¶](https://ondemand.cluster.france-bioinformatique.fr/node/cpu-node-23.ifb.local/52978/lab/tree/shared/projects/2412_fairbioinfo/usecase/Deseq2.ipynb#Deseq2)

### Summary

`DESeq2` is a widely used bioinformatics tool for differential gene expression analysis from RNA sequencing (RNA-seq) data. Here's a concise overview of its functionality:

1.  **Differential Expression Analysis:** DESeq2 performs statistical analysis to identify genes that are differentially expressed between experimental conditions or sample groups.

2.  **Input Data:** Accepts count data derived from RNA-seq experiments, typically in the form of read counts mapped to genomic features such as genes or transcripts.

3.  **Normalization:** DESeq2 employs normalization techniques to account for differences in sequencing depth and other technical biases across samples, ensuring accurate comparisons between conditions.

4.  **Statistical Modeling:** Utilizes negative binomial distribution-based models to estimate variance and test for differential expression, taking into account sample variability and experimental design factors.

5.  **Multiple Testing Correction:** Corrects for multiple hypothesis testing using methods such as the Benjamini-Hochberg procedure to control the false discovery rate (FDR) and reduce the likelihood of spurious results.

6.  **Visualization:** Provides functions for visualizing expression data, including heatmaps, volcano plots, and principal component analysis (PCA) plots, facilitating interpretation and exploration of results.

7.  **Annotation:** Allows for annotation of differentially expressed genes with biological metadata such as gene symbols, functional annotations, and pathway information, aiding in biological interpretation.

8.  **Integration with Workflows:** DESeq2 can be seamlessly integrated into bioinformatics pipelines and workflows, serving as a fundamental tool for gene expression analysis in diverse biological studies.

In summary, `DESeq2` is a powerful and widely adopted tool for differential gene expression analysis from RNA-seq data, providing robust statistical methods and comprehensive functionalities for biological interpretation.

### Code

```{r}
# Get Data 
data = read.table(file = "deseq2_demo.tsv",  header = TRUE, row.names = 1, sep="\t")
data$Gene = rownames(data)
head(data)
```

## Volcano plot en ggplot

```{r}
library(ggplot2)

pvalue_threshold <- 0.05

p <- ggplot(data, aes(x = log2FoldChange, y = -log10(pvalue))) +
  geom_point(color = ifelse(data$pvalue < pvalue_threshold, "red", "black")) +
  scale_color_manual(values = c("black", "red")) +
  geom_text(data = subset(data, pvalue < pvalue_threshold), aes(label = Gene), vjust = 1) +
  labs(x = "Log2FoldChange", y = "-log10(p-value)") +
  theme_minimal()

p

```

## Volcano plot en Plotly

```{r}
library(plotly)
p <- ggplotly(p)
p
```
